Informative Oceanographic Sampling using Excursion Probabilities for Multivariate Random Fields

### 2019-09-06 main_new.tex
updates

Our objective is to characterize a river plume, focusing on spatial
separation of cold freshwater from the warmer saline waters of the
fjord. ESs and EPs are useful starting points to measure the ability
to classify water masses, and the goal of experimental sampling is to
improve this characterization, as quantified via changes in the EPs.

We consider a lateral domain, near sea surface, with locations
$\bx \in \mathcal{M} \subset \mathbf{R^2}$. The two variables of
interest, temperature and salinity, are defined as random processes;
$\xi_a(\bx)$ denotes the temperature (in $^o C$) and $\xi_b(\bx)$
denotes the salinity (in $mg/l$) at location $\bx$.  We let $t_a$ be a
threshold for temperature and $t_b$ a threshold for salinity, the
bivariate ES is then defined by
 \begin{equation}\label{ES}
     \mbox{ES} = \{\bx \in \mathcal{M} | \xi_a(\bx) \leq t_a,\xi_b(\bx) \leq t_b\}.
 \end{equation}
 The excursions could also be defined as larger than a threshold, or
 within a boundary of limits for the two variables. Calculations and
 interpretations will be similar in these different cases.

The ES holds random indicator variables, while the associated EPs are 
\begin{equation}\label{eq:prob}
 p(\bx) = P(\xi_a(\bx) \leq t_a, \xi_b(\bx) \leq t_b), \hspace{3mm} \bx \in \mathcal{M}.
\end{equation}
If the EPs in Eq. (\ref{eq:prob}) are
close to $1$ or $0$, it is easy to classify water masses; it is
challenging if the probabilities are close to $0.5$. 
The integrated Bernoulli variance (IBV), see \cite{bect2019}, is here defined as 
\begin{equation}\label{measVAR}
    V = \int_{\bx \in \mathcal{M}} p(\bx) \left(1-p(\bx)\right) d\bx.
\end{equation}
The measure is dominated by locations with EPs close to $0.5$.

If prior knowledge strongly indicates cold freshwater in a part of the domain, it is easy to classify, and there is likely little value in exploring that part much. Similarly with strong prior knowledge about warm saltwater. 
Data can be gathered at selected locations of the domain $\mathcal{M}$, to improve classification skill. We denote temperature measurements by $y_{a}(\bx_i)$ and salinity measurements by $y_{b}(\bx_i)$, at a sampling location $\bx_i \in \mathcal{M}$. Given a set of measurements, denoted $\by$, the conditional EPs are
\begin{equation}\label{eq:post_ep}
 p(\bx;\by) = P(\xi_a(\bx) \leq t_a, \xi_b(\bx) \leq t_b |\by). 
\end{equation}

Our goal is to construct designs for the data gathering, that are expected to improve the mapping of the water masses. 
When designing AUV sampling designs, it is hence natural to evaluate the expected reduction in the IBV, as provided by the survey data. In this way, we take the expectation with respect to the planned measurements $\by$;
\begin{equation}\label{sur}
    V_{\mbox{upd}} = \int_{\bx \in \mathcal{M}} E_{\by} \left\{ p(\bx;\by)\left( 1-p(\bx;\by)\right) \right\} d\bx, 
\end{equation}
where the probability is defined in Eq. (\ref{eq:post_ep}).
Considering a set $\mathcal{J}$ of possible designs for an AUV survey. A criterion for the selection is
\begin{equation}\label{crit}
    j^* = \mbox{argmin}_{j \in \mathcal{J}} V_{\mbox{upd}}(j),
\end{equation}
where the criterion $V_{\mbox{upd}}(j)$ in Eq. (\ref{sur}) is computed for each of the possible designs.

Recent works in statistical literature on ESs and EPs connected to
spatial statistics include
\cite{picheny2010,french2013spatio,bolin2015excursion,french2016credible}. Our focus
is on ES as defined in Eq. (\ref{ES}), and the
uncertainty reduction achieved by sampling as in Eq. (\ref{sur}).  In this sense
our work is similar to \cite{bect2012}, \cite{chevalier2014fast} and
\cite{azzimonti2016quantifying} who describe analytical results for
uncertainty reduction in ESs for univariate processes.  
A major contribution of our work is to derive closed-form results for this design criteria in the situation with multivariate GPs.

>>>>

Our objective is to characterize a river plume, focusing on spatial
separation of cold freshwater from the warmer saline waters of the
fjord. Excursion sets (ES) and excursion probabilities (EP) are useful
starting points to measure the ability to classify water masses. The
goal of experimental sampling is to improve this characterization, as
quantified via changes in the EPs.

We consider a lateral domain, near the sea surface, with locations
$\bx \in \mathcal{M} \subset \mathbf{R^2}$. The two variables of
interest, temperature and salinity, are defined as random processes;
$\xi_a(\bx)$ denotes the temperature (in $^o C$) and $\xi_b(\bx)$
denotes the salinity (in $mg/l$) at location $\bx$.  We let $t_a$ be a
threshold for temperature and $t_b$ a threshold for salinity, the
bivariate ES is then defined by

\begin{equation}\label{ES}
     \mbox{ES} = \{\bx \in \mathcal{M} | \xi_a(\bx) \leq t_a,\xi_b(\bx) \leq t_b\}.
\end{equation}
The excursions could also be defined as larger than a threshold, or
within a boundary of limits for the two variables; calculations and
interpretations will be similar in these cases.

The ES holds random indicator variables, while the associated EPs are 
\begin{equation}\label{eq:prob}
  p(\bx) = P(\xi_a(\bx) \leq t_a, \xi_b(\bx) \leq t_b), \hspace{3mm} \bx \in \mathcal{M}.
\end{equation}
If the EPs in Eq. (\ref{eq:prob}) are
close to $1$ or $0$, it is easy to classify water masses; it is
challenging if the probabilities are close to $0.5$. 
The integrated Bernoulli variance (IBV) \cite{bect2019}, is defined as 
\begin{equation}\label{measVAR}
    V = \int_{\bx \in \mathcal{M}} p(\bx) \left(1-p(\bx)\right) d\bx.
\end{equation}
The measure is dominated by locations with EPs close to $0.5$.
\kc{This sentence is a bit strange; just about, you state that EPs
  close to 0.5 can be problematic since they do not categorize the
  water masses well. And here you state that there is a strong
  dominance of the average. So now you should say how the biasing can
  change to make ``V'' more effective. Otherwise this sentence is
  dangling.}

If prior knowledge provides a strong indication of cold freshwater
(for instance) in a part of the domain, it is easy to classify, and
there is likely little value in exploring those locations. Data can be
gathered at selected locations of the domain $\mathcal{M}$, to improve
classification skill. We denote temperature measurements by
$y_{a}(\bx_i)$ and salinity measurements by $y_{b}(\bx_i)$, at a
sampling location $\bx_i \in \mathcal{M}$. Given a set of
measurements, denoted $\by$, the conditional EPs are
\begin{equation}\label{eq:post_ep}
 p(\bx;\by) = P(\xi_a(\bx) \leq t_a, \xi_b(\bx) \leq t_b |\by). 
\end{equation}

Our goal is to construct designs for data gathering, that are expected
to improve the mapping and separation of water masses. When designing
AUV sampling designs, it is therefore natural to evaluate the expected
reduction in the IBV, as provided by the survey data. In this way, we
take the expectation with respect to the planned measurements $\by$;
\begin{equation}\label{sur}
    V_{\mbox{upd}} = \int_{\bx \in \mathcal{M}} E_{\by} \left\{ p(\bx;\by)\left( 1-p(\bx;\by)\right) \right\} d\bx, 
\end{equation}
where the probability is defined in Eq. (\ref{eq:post_ep}).
Considering a set $\mathcal{J}$ of possible designs for an AUV survey. A criterion for the selection is
\begin{equation}\label{crit}
    j^* = \mbox{argmin}_{j \in \mathcal{J}} V_{\mbox{upd}}(j),
\end{equation}
where the criterion $V_{\mbox{upd}}(j)$ in Eq. (\ref{sur}) is computed for each of the possible designs.

Recent work in literature on ESs and EPs connected to spatial
statistics include
\cite{picheny2010,french2013spatio,bolin2015excursion,french2016credible}.
Our focus is on ES as defined in Eq. (\ref{ES}), and the uncertainty
reduction achieved by sampling as in Eq. (\ref{sur}). In this sense
our work is similar to \cite{bect2012}, \cite{chevalier2014fast} and
\cite{azzimonti2016quantifying} who describe analytical results for
uncertainty reduction in ESs for univariate processes. A major
contribution of our work is to derive closed-form results for this
design criteria in the situation with multivariate GPs.
